|

### 2022-09-05 paper_comparison/paper_comparison/paper_comparison.Rmd
polish the residual departure

<!-- problem: residual plot diagnostics -->
<!-- conventional test: too sensitive -->

@@

<!-- week 1 -->
<!-- 1. residual plot for model diagnostics -->

<!-- a. residual is widely used -->
<!-- b. what are the types of residual plots -->
<!-- c. comparison -->

<!-- week 2 -->
<!-- 2. conventional test: F, BP -->
<!-- 3. visual test: lineup, theory -->

<!-- week 4 -->
<!-- desc of experiment: -->
<!-- 1. simulation setup -->
<!-- 2. experimental design -->
<!-- 3. result -->

<!-- week 3 -->
<!-- comparison of conventional tests: -->
<!-- 1. power (visual test vs. conventional test) -->
<!-- (visual test most different one (everything test, any departure)) -->
<!-- plot figure in a paper, desc, exp -->
<!-- 2. investigate the difference (gap), give examples -->
<!-- 3. conventional is too sensitive -->
<!-- 4. make conventional less sensitive (vary alpha) -->

<!-- last week -->
<!-- conclusion: -->
<!-- 1. too sensitive, visual test is needed/preferable -->
<!-- 2. visual test is infeasible in large scale (expensive) -->
<!-- 3. future work (role of computer vision) -->

@@

<!--Regression diagnostics is an essential step in regression analysis which is a field of study with at least a hundred years of history. The diagnostic procedure conventionally involves evaluating the fitness of the proposed model, detecting the presence of influential observations and outliers, checking the validity of model assumptions and many more. In terms of diagnostic techniques, data plots, hypothesis testing, and summary statistics are vital tools for a systematic and detailed examination of the regression model [@mansfield1987diagnostic].

Many of those regression diagnostic methods and procedures are mature and well-established in books first published in the twentieth century, such as @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals. Regardless of the level of difficulty, one will find the importance and usefulness of diagnostic plots being emphasized in those books repeatedly. Checking diagnostic plots is also the recommended starting point for validating model assumptions such as normality, homoscedasticity and linearity [@anscombe_examination_1963]. 
-->

# Background

@@

There are many departures from a good residual plot, but we will consider only three types of them in this paper, which are also the most commonly checked departures for a residual plot. The first one is non-linearity. Any unnatural non-linear curve presented in the residual plot could be considered an indicative of non-linearity. At the top-right of Figure [ref here], there is an example residual plot of this type of departure. The second type of departure  we will consider is heterskedasticity. The usual pattern of this type of departure is the inconsistent distribution of the residuals for different x values. Visually, it usually results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure [ref here], or the "triangle" shape where the smallest variance occurs at the edges of the x-axis. The last type of departure is non-normality. This departure is usually harder to be detected from the residuals plot compared to non-linearity and heterskedasticity. In most the case, a quantile-quantile plot can do a better job. Depending on the actual distribution of the residuals, different visual patterns will be discovered. If the distribution is skewed, one will see some outliers presented on only one side of the zero line as shown in the bottom-right of Figure [ref here].

@@

<!-- ![Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009]. Visible features in a plot are viewed as a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$, and any visual discoveries that are inconsistent with the null hypothesis are treated as evidence against the null. For regression diagnostics, the null hypothesis would be the assumed model, and visual discoveries could be any visual features in favour of any alternatives. \label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341}
-->

In visual inference, a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ is defined, where $\boldsymbol{\mathrm{y}}$ is the data and $I$ is a set of all possible visual features. @buja_statistical_2009 described each of the test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})$ as a measurement of the degree of presence of a visual feature. Alternatively, @majumder_validation_2013 avoids the use of visual features and defined the visual statistics $T(.)$ as a mapping from a dataset to a data plot. Both definitions of visual test statistics are valid, but in the rest of the paper the first definition will be used as it covers some details needed by the following discussion. A visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries [@buja_statistical_2009]. For regression diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis. The same regression model can be rejected by many reasons with residual plot, including non-linearity and heteroskedasticity as shown in Figure \ref{fig:residual-plot-cubic-heter}.

@@

\begin{align} \label{eq:dirichlet-multinomial}\begin{split}
\boldsymbol{\theta}&|\alpha \sim Dirichlet(\alpha)\\
(c_i,...,c_m)&|\boldsymbol{\theta} \sim Multinomial(K, \boldsymbol{\theta}).
\end{split}\end{align}

Since the p-value calculation only needs to concern the number of times the actual data plot being selected denoted by $C_i$, they showed the model can be simplified to a beta-binomial mixture distribution

\begin{align} \label{eq:beta-binomial}\begin{split}
\theta_i&|\alpha \sim Beta(\alpha, (m-1)\alpha)\\
C_i&|\theta_i \sim Binomial(K, \theta_i).
\end{split}\end{align}

Thus, the visual p-value followed by the beta-binomial model is given as

\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \sum_{x=c_i}^{K}{{K}\choose{x}}\frac{1}{B(\alpha, (m-1)\alpha)}B(x + \alpha, K - x + (m - 1)\alpha),
\end{equation}

where $B(.)$ is the beta function defined as

\begin{equation} \label{eq:betafunction}
B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0. 
\end{equation}

Note that the use of Equation (\ref{eq:pvalue-beta-binomial}) requires the estimation of $\hat{\alpha}$, which largely depends on the null model, the type of the plot and other aesthetic features. They suggested to estimate $\hat{\alpha}$ visually based on the selections of null plots of the experimental data, or to estimate $\hat{\alpha}$ numerically based on several additional Rorschach lineups, which is a type of lineup containing only null plots. However, when the number of null models are large, it could be expensive to manually estimate each $\alpha$ or include additional Rorschach lineups in the experiment.

@@

\begin{equation} \label{eq:p-value-multi}
\text{p-value} = \frac{\#draws~that~the~actual~data~plota~being~selected~more~than~c_i~times}{\#simulation}.
\end{equation}

<!-- ### Effectiveness of visual inference in regression diagnostics -->

@@

<!-- $$g(p_{li}) = W_{li}\delta + Z_{li}\tau_{li},$$ -->
<!-- where $g(.)$ is the logit link function $g(p) = log(p)  - log(1-p)$; $0 \leq p \leq 1$. $W_{li}$, $1 \leq i \leq K$, $1 \leq l \leq L$, is the covariate matrix including lineup-specific elements and demographic information of individuals, and $\delta$ is a vector of parameters. $Z$ is the random effects matrix, and $\tau$ is a vector of variables follow $N(\boldsymbol{0},\sigma_{\tau}\boldsymbol{I}_{KL\times KL})$.  -->

@@

\begin{align} \label{eq:nonlinearity-model}
\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1), \\
\boldsymbol{z} = g(\boldsymbol{z}_{raw}, 1), \\
\boldsymbol{z}_{raw} = He_j(g(\boldsymbol{z}, 2)),
\end{align}

where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vector of size $n$,  $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $\{-k, k\}$ defined as

\begin{equation} \label{eq:scaling-function}
g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x})) \times 2k - k, \quad \text{for} \quad k > 0. 
\end{equation}

The null regression model used to fit the realizations generated by the above model is formulated as:

\begin{equation} \label{eq:scaling-function}
\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
\end{equation}

where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.

@@

\begin{align} \label{eq:heter-model}
\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
\end{align}

where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vector of size $n$.

@@

\begin{equation} 
\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
\end{equation}

where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.

@@

\begin{equation} \label{eq:effect-size-ex1}
E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b,
\end{equation}

where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.

@@

\begin{equation} \label{eq:logistic-regression-1-1}
Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})),
\end{equation}

where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.

@@

Figure \ref{fig:power-overview} shows an overview of estimated power of visual test against natural logarithm of the effect with comparison to the power of an exact test - F-test, and the power of two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions. In overall, the power of all four tests increases as the effect becomes larger. The power curve of F-test climbs aggressively from 25\% to around 90\% as $log_e(E)$ increases from 0 to 2, while others respond inactively to the change of effect and remain lower than 25\% throughout the period, showing that as an exact test, the F-test is relatively more sensitive to the type of model defects that being considered. The power of visual test arises steadily and nearly linearly to around 90\% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make noticeable impact on the degree of the presence of the designed visual features. Other two inappropriate conventional tests shows improvement at the same time but at a lower rate. This coincides the point made by @cook1982residuals mentioned in \ref{hypothesis-testing} that residual-based tests for a specific type of model defect are sensitive to other types of model defects. At $log_e(E) = 6$, the power curve of F-test reaches almost 100% followed by the visual test by a small margin. The power of Breusch–Pagan test and Shapiro–Wilk test reach around 75% and 63% respectively.

@@

<!-- # time taken vs order of the plot -->
<!-- # accuracy vs order of the plot -->
<!-- # num of selection vs order of the plot -->
<!-- # order of the plot? -->
<!-- # summary of the demographic information -->

@@

<!-- 1. power (visual test vs. conventional test) -->
<!-- (visual test most different one (everything test, any departure)) -->
<!-- plot figure in a paper, desc, exp -->
<!-- 2. investigate the difference (gap), give examples -->
<!-- 3. conventional is too sensitive -->
<!-- 4. make conventional less sensitive (vary alpha) -->

@@

<!-- <!-- sample 1, change values of a and b, plot the data  --> -->

@@

<!-- ```{r power-vs-log-effect-size} -->
<!-- polynomials %>% -->
<!--   group_by(lineup_id) %>% -->
<!--   summarise(across(c(type:p_value), first)) %>% -->
<!--   filter(lineup_id < 577) %>% -->
<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
<!--   ggplot() + -->
<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   xlab("Natural logarithm of effect size") + -->
<!--   ylab("Power") -->
<!-- ``` -->

<!-- ```{r power-vs-log-effect-size-given-x-dist} -->
<!-- polynomials %>% -->
<!--   group_by(lineup_id) %>% -->
<!--   summarise(across(c(type:p_value), first)) %>% -->
<!--   filter(lineup_id < 577) %>% -->
<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
<!--   ggplot() + -->
<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   facet_wrap(~x_dist) + -->
<!--   xlab("Natural logarithm of effect size") + -->
<!--   ylab("Power") + -->
<!--   ggtitle("Power comparison conditional on the distribution of X") -->
<!-- ``` -->

<!-- ```{r power-of-visual-test-given-x-dist} -->
<!-- polynomials %>% -->
<!--   group_by(lineup_id) %>% -->
<!--   summarise(across(c(type:p_value), first)) %>% -->
<!--   filter(lineup_id < 577) %>% -->
<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
<!--   ggplot() + -->
<!--   geom_smooth(aes(log(effect_size), reject, col = x_dist), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   xlab("Natural logarithm of effect size") + -->
<!--   ylab("Power") + -->
<!--   ggtitle("Power of visual test conditional on the distribution of X") -->
<!-- ``` -->

<!-- ```{r power-vs-log-effect-size-given-shape} -->
<!-- polynomials %>% -->
<!--   group_by(lineup_id) %>% -->
<!--   summarise(across(c(type:p_value), first)) %>% -->
<!--   filter(lineup_id < 577) %>% -->
<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
<!--   ggplot() + -->
<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   facet_wrap(~shape) + -->
<!--   xlab("Natural logarithm of effect size") + -->
<!--   ylab("Power") + -->
<!--   ggtitle("Power comparison conditional on the shape of the Hermite polynomials") -->
<!-- ``` -->

<!-- ```{r power-of-visual-test-given-shape} -->
<!-- polynomials %>% -->
<!--   group_by(lineup_id) %>% -->
<!--   summarise(across(c(type:p_value), first)) %>% -->
<!--   filter(lineup_id < 577) %>% -->
<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01)), shape = factor(shape)) %>% -->
<!--   ggplot() + -->
<!--   geom_smooth(aes(log(effect_size), reject, col = shape), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   xlab("Natural logarithm of effect size") + -->
<!--   ylab("Power") + -->
<!--   ggtitle("Power of visual test conditional on the shape of the Hermite polynomials") -->
<!-- ``` -->

<!-- ```{r power-vs-log-effect-size-given-number-of-observations} -->
<!-- polynomials %>% -->
<!--   group_by(lineup_id) %>% -->
<!--   summarise(across(c(type:p_value), first)) %>% -->
<!--   filter(lineup_id < 577) %>% -->
<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->
<!--   ggplot() + -->
<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->
<!--   facet_wrap(~n) + -->
<!--   xlab("Natural logarithm of effect size") + -->
<!--   ylab("Power") + -->
<!--   ggtitle("Power comparison conditional on the number of observations") -->
<!-- ``` -->

>>>>

<!-- problem: residual plot diagnostics -->

<!-- conventional test: too sensitive -->

@@

<!-- week 1 -->

<!-- 1. residual plot for model diagnostics -->

<!-- a. residual is widely used -->

<!-- b. what are the types of residual plots -->

<!-- c. comparison -->

<!-- week 2 -->

<!-- 2. conventional test: F, BP -->

<!-- 3. visual test: lineup, theory -->

<!-- week 4 -->

<!-- desc of experiment: -->

<!-- 1. simulation setup -->

<!-- 2. experimental design -->

<!-- 3. result -->

<!-- week 3 -->

<!-- comparison of conventional tests: -->

<!-- 1. power (visual test vs. conventional test) -->

<!-- (visual test most different one (everything test, any departure)) -->

<!-- plot figure in a paper, desc, exp -->

<!-- 2. investigate the difference (gap), give examples -->

<!-- 3. conventional is too sensitive -->

<!-- 4. make conventional less sensitive (vary alpha) -->

<!-- last week -->

<!-- conclusion: -->

<!-- 1. too sensitive, visual test is needed/preferable -->

<!-- 2. visual test is infeasible in large scale (expensive) -->

<!-- 3. future work (role of computer vision) -->

@@

```{=html}
<!--Regression diagnostics is an essential step in regression analysis which is a field of study with at least a hundred years of history. The diagnostic procedure conventionally involves evaluating the fitness of the proposed model, detecting the presence of influential observations and outliers, checking the validity of model assumptions and many more. In terms of diagnostic techniques, data plots, hypothesis testing, and summary statistics are vital tools for a systematic and detailed examination of the regression model [@mansfield1987diagnostic].

Many of those regression diagnostic methods and procedures are mature and well-established in books first published in the twentieth century, such as @draper1998applied, @montgomery1982introduction, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals. Regardless of the level of difficulty, one will find the importance and usefulness of diagnostic plots being emphasized in those books repeatedly. Checking diagnostic plots is also the recommended starting point for validating model assumptions such as normality, homoscedasticity and linearity [@anscombe_examination_1963]. 
-->
```
# Background

@@

There are many types of departures from a good residual plot. Three of them will be considered in this paper, which are also the most commonly checked departures for a residual plot. The first one is non-linearity. Any unnatural and non-linear functional form of residuals on fitted values presented in the residual plot could be considered as an indicative of non-linearity. At the top-right of Figure [ref here], there is an example residual plot of this type of departure. The second type of departure that will be considered is heterskedasticity. The usual pattern is the inconsistent distribution of the residuals for different x values. Visually, it sometimes results in the so-called "butterfly" shape as shown in the bottom-left panel of Figure [ref here], or the "triangle" shape where the smallest variance occurs at the edges of the x-axis. The last type of departure we will consider is non-normality. This departure is usually harder to be detected from the residuals plot compared to non-linearity and heterskedasticity. In most cases, a quantile-quantile plot can do a better job. Depending on the actual distribution of the residuals, different visual patterns will be discovered. Readers of the residuals plot will notice the data points do not distribute noramlly on the y-axis. For example, if the distribution is skewed, one will see the "tail" of the distribution as some outliers presented on only one side of the zero line as shown in the bottom-right of Figure [ref here].

@@

```{=html}
<!-- ![Parallelism between multiple quantitative testing and visual discovery [@buja_statistical_2009]. Visible features in a plot are viewed as a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$, and any visual discoveries that are inconsistent with the null hypothesis are treated as evidence against the null. For regression diagnostics, the null hypothesis would be the assumed model, and visual discoveries could be any visual features in favour of any alternatives. \label{fig:parallelism}](figures/rsta2009012001.jpg){width=450 height=341}
-->
```
In visual inference, a collection of test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})~(i \in I)$ is defined, where $\boldsymbol{\mathrm{y}}$ is the data and $I$ is a set of all possible visual features. @buja_statistical_2009 described each of the test statistics $T^{(i)}(\boldsymbol{\mathrm{y}})$ as a measurement of the degree of presence of a visual feature. Alternatively, @majumder_validation_2013 avoids the use of visual features and defined the visual statistics $T(.)$ as a mapping from a dataset to a data plot. Both definitions of visual test statistics are valid, but in the rest of the paper the first definition will be used as it covers some details needed by the following discussion. A visual discovery is defined as a rejection of a null hypothesis, and the same null hypothesis can be rejected by many different visual discoveries [@buja_statistical_2009]. For regression diagnostics, the null hypothesis would be the assumed model, while the visual discoveries would be any findings that are inconsistent with the null hypothesis. The same regression model can be rejected by many reasons with residual plot, including non-linearity and heteroskedasticity as shown in Figure \ref{fig:residual-plot-cubic-heter}.

@@

```{=tex}
\begin{align} \label{eq:dirichlet-multinomial}\begin{split}
\boldsymbol{\theta}&|\alpha \sim Dirichlet(\alpha)\\
(c_i,...,c_m)&|\boldsymbol{\theta} \sim Multinomial(K, \boldsymbol{\theta}).
\end{split}\end{align}
```
Since the p-value calculation only needs to concern the number of times the actual data plot being selected denoted by $C_i$, they showed the model can be simplified to a beta-binomial mixture distribution

```{=tex}
\begin{align} \label{eq:beta-binomial}\begin{split}
\theta_i&|\alpha \sim Beta(\alpha, (m-1)\alpha)\\
C_i&|\theta_i \sim Binomial(K, \theta_i).
\end{split}\end{align}
```
Thus, the visual p-value followed by the beta-binomial model is given as

```{=tex}
\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \sum_{x=c_i}^{K}{{K}\choose{x}}\frac{1}{B(\alpha, (m-1)\alpha)}B(x + \alpha, K - x + (m - 1)\alpha),
\end{equation}
```
where $B(.)$ is the beta function defined as

```{=tex}
\begin{equation} \label{eq:betafunction}
B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0. 
\end{equation}
```
Note that the use of Equation (\ref{eq:pvalue-beta-binomial}) requires the estimation of $\hat{\alpha}$, which largely depends on the null model, the type of the plot and other aesthetic features. They suggested to estimate $\hat{\alpha}$ visually based on the selections of null plots of the experimental data, or to estimate $\hat{\alpha}$ numerically based on several additional Rorschach lineups, which is a type of lineup containing only null plots. However, when the number of null models are large, it could be expensive to manually estimate each $\alpha$ or include additional Rorschach lineups in the experiment.

@@

```{=tex}
\begin{equation} \label{eq:p-value-multi}
\text{p-value} = \frac{\#draws~that~the~actual~data~plota~being~selected~more~than~c_i~times}{\#simulation}.
\end{equation}
```
<!-- ### Effectiveness of visual inference in regression diagnostics -->

@@

<!-- $$g(p_{li}) = W_{li}\delta + Z_{li}\tau_{li},$$ -->

<!-- where $g(.)$ is the logit link function $g(p) = log(p)  - log(1-p)$; $0 \leq p \leq 1$. $W_{li}$, $1 \leq i \leq K$, $1 \leq l \leq L$, is the covariate matrix including lineup-specific elements and demographic information of individuals, and $\delta$ is a vector of parameters. $Z$ is the random effects matrix, and $\tau$ is a vector of variables follow $N(\boldsymbol{0},\sigma_{\tau}\boldsymbol{I}_{KL\times KL})$.  -->

@@

```{=tex}
\begin{align} \label{eq:nonlinearity-model}
\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{z} + \boldsymbol{\varepsilon},\\
\boldsymbol{x} = g(\boldsymbol{x}_{raw}, 1), \\
\boldsymbol{z} = g(\boldsymbol{z}_{raw}, 1), \\
\boldsymbol{z}_{raw} = He_j(g(\boldsymbol{z}, 2)),
\end{align}
```
where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$, $\boldsymbol{x}_{raw}$, $\boldsymbol{z}_{raw}$ are vector of size $n$, $He_{j}(.)$ is the $j$th-order probabilist's Hermite polynomials, $\varepsilon \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$, and $g(\boldsymbol{x}, k)$ is a scaling function to enforce the support of the random vector to be $\{-k, k\}$ defined as

```{=tex}
\begin{equation} \label{eq:scaling-function}
g(\boldsymbol{x}, k) = (\boldsymbol{x} - min(\boldsymbol{x}))/max(\boldsymbol{x} - min(\boldsymbol{x})) \times 2k - k, \quad \text{for} \quad k > 0. 
\end{equation}
```
The null regression model used to fit the realizations generated by the above model is formulated as:

```{=tex}
\begin{equation} \label{eq:scaling-function}
\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
\end{equation}
```
where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.

@@

```{=tex}
\begin{align} \label{eq:heter-model}
\boldsymbol{y} = 1 + \boldsymbol{x} + \boldsymbol{\varepsilon},\\
\boldsymbol{\varepsilon} \sim N(\boldsymbol{0}, 1 + 2 - |a| b (\boldsymbol{x} - a)^2 \boldsymbol{I}), \\
\end{align}
```
where $\boldsymbol{y}$, $\boldsymbol{x}$, $\boldsymbol{\varepsilon}$ are vector of size $n$.

@@

```{=tex}
\begin{equation} 
\boldsymbol{y} = \beta_0 + \beta_1 \boldsymbol{x} + \boldsymbol{u},
\end{equation}
```
where $\boldsymbol{u} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{I}_n)$.

@@

```{=tex}
\begin{equation} \label{eq:effect-size-ex1}
E = \frac{1}{2\sigma^2}\boldsymbol{X}_b'\boldsymbol{R}_a'(diag(\boldsymbol{R}_a))^{-1}\boldsymbol{R}_a\boldsymbol{X}_b,
\end{equation}
```
where $diag(.)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.

@@

```{=tex}
\begin{equation} \label{eq:logistic-regression-1-1}
Pr(\text{reject}~H_0|H_1,E) = \Lambda(\beta_0 + \beta_1 log_e(\boldsymbol{E})),
\end{equation}
```
where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$.

@@

Figure \ref{fig:power-overview} shows an overview of estimated power of visual test against natural logarithm of the effect with comparison to the power of an exact test - F-test, and the power of two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions. In overall, the power of all four tests increases as the effect becomes larger. The power curve of F-test climbs aggressively from 25% to around 90% as $log_e(E)$ increases from 0 to 2, while others respond inactively to the change of effect and remain lower than 25% throughout the period, showing that as an exact test, the F-test is relatively more sensitive to the type of model defects that being considered. The power of visual test arises steadily and nearly linearly to around 90% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make noticeable impact on the degree of the presence of the designed visual features. Other two inappropriate conventional tests shows improvement at the same time but at a lower rate. This coincides the point made by @cook1982residuals mentioned in \ref{hypothesis-testing} that residual-based tests for a specific type of model defect are sensitive to other types of model defects. At $log_e(E) = 6$, the power curve of F-test reaches almost 100% followed by the visual test by a small margin. The power of Breusch--Pagan test and Shapiro--Wilk test reach around 75% and 63% respectively.

@@

<!-- # time taken vs order of the plot -->

<!-- # accuracy vs order of the plot -->

<!-- # num of selection vs order of the plot -->

<!-- # order of the plot? -->

<!-- # summary of the demographic information -->

@@

<!-- 1. power (visual test vs. conventional test) -->

<!-- (visual test most different one (everything test, any departure)) -->

<!-- plot figure in a paper, desc, exp -->

<!-- 2. investigate the difference (gap), give examples -->

<!-- 3. conventional is too sensitive -->

<!-- 4. make conventional less sensitive (vary alpha) -->

@@

<!-- <!-- sample 1, change values of a and b, plot the data  -->

--\>

@@

<!-- ```{r power-vs-log-effect-size} -->

<!-- polynomials %>% -->

<!--   group_by(lineup_id) %>% -->

<!--   summarise(across(c(type:p_value), first)) %>% -->

<!--   filter(lineup_id < 577) %>% -->

<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->

<!--   ggplot() + -->

<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->

<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->

<!--   xlab("Natural logarithm of effect size") + -->

<!--   ylab("Power") -->

<!-- ``` -->

<!-- ```{r power-vs-log-effect-size-given-x-dist} -->

<!-- polynomials %>% -->

<!--   group_by(lineup_id) %>% -->

<!--   summarise(across(c(type:p_value), first)) %>% -->

<!--   filter(lineup_id < 577) %>% -->

<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->

<!--   ggplot() + -->

<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->

<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->

<!--   facet_wrap(~x_dist) + -->

<!--   xlab("Natural logarithm of effect size") + -->

<!--   ylab("Power") + -->

<!--   ggtitle("Power comparison conditional on the distribution of X") -->

<!-- ``` -->

<!-- ```{r power-of-visual-test-given-x-dist} -->

<!-- polynomials %>% -->

<!--   group_by(lineup_id) %>% -->

<!--   summarise(across(c(type:p_value), first)) %>% -->

<!--   filter(lineup_id < 577) %>% -->

<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->

<!--   ggplot() + -->

<!--   geom_smooth(aes(log(effect_size), reject, col = x_dist), method = "glm", method.args = list(family = binomial), se = FALSE) + -->

<!--   xlab("Natural logarithm of effect size") + -->

<!--   ylab("Power") + -->

<!--   ggtitle("Power of visual test conditional on the distribution of X") -->

<!-- ``` -->

<!-- ```{r power-vs-log-effect-size-given-shape} -->

<!-- polynomials %>% -->

<!--   group_by(lineup_id) %>% -->

<!--   summarise(across(c(type:p_value), first)) %>% -->

<!--   filter(lineup_id < 577) %>% -->

<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->

<!--   ggplot() + -->

<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->

<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->

<!--   facet_wrap(~shape) + -->

<!--   xlab("Natural logarithm of effect size") + -->

<!--   ylab("Power") + -->

<!--   ggtitle("Power comparison conditional on the shape of the Hermite polynomials") -->

<!-- ``` -->

<!-- ```{r power-of-visual-test-given-shape} -->

<!-- polynomials %>% -->

<!--   group_by(lineup_id) %>% -->

<!--   summarise(across(c(type:p_value), first)) %>% -->

<!--   filter(lineup_id < 577) %>% -->

<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01)), shape = factor(shape)) %>% -->

<!--   ggplot() + -->

<!--   geom_smooth(aes(log(effect_size), reject, col = shape), method = "glm", method.args = list(family = binomial), se = FALSE) + -->

<!--   xlab("Natural logarithm of effect size") + -->

<!--   ylab("Power") + -->

<!--   ggtitle("Power of visual test conditional on the shape of the Hermite polynomials") -->

<!-- ``` -->

<!-- ```{r power-vs-log-effect-size-given-number-of-observations} -->

<!-- polynomials %>% -->

<!--   group_by(lineup_id) %>% -->

<!--   summarise(across(c(type:p_value), first)) %>% -->

<!--   filter(lineup_id < 577) %>% -->

<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01))) %>% -->

<!--   ggplot() + -->

<!--   geom_smooth(aes(log(effect_size), reject, col = "Visual test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->

<!--   geom_smooth(data = conv_result, aes(log(effect_size), reject, col = "Conventional test"), method = "glm", method.args = list(family = binomial), se = FALSE) + -->

<!--   facet_wrap(~n) + -->

<!--   xlab("Natural logarithm of effect size") + -->

<!--   ylab("Power") + -->

<!--   ggtitle("Power comparison conditional on the number of observations") -->

<!-- ``` -->

<!-- ```{r power-of-visual-test-given-number-of-observations} -->

<!-- polynomials %>% -->

<!--   group_by(lineup_id) %>% -->

<!--   summarise(across(c(type:p_value), first)) %>% -->

<!--   filter(lineup_id < 577) %>% -->

<!--   mutate(reject = as.numeric(eval_p_value(p_value, tol = 0.01)), n = factor(n)) %>% -->

<!--   ggplot() + -->

<!--   geom_smooth(aes(log(effect_size), reject, col = n), method = "glm", method.args = list(family = binomial), se = FALSE) + -->

<!--   xlab("Natural logarithm of effect size") + -->

<!--   ylab("Power") + -->

<!--   ggtitle("Power of visual test conditional on the number of observations") -->

<!-- ``` -->

<!-- ```{r} -->

<!-- polynomials %>% -->

<!--   group_by(lineup_id) %>% -->
